Level-Set method on \texttt{p4est}

### 2014-11-13 paper/tex/parallel_algorithms.tex
Fixed minor issues in .tex. Added draft.pdf
more edits
added interpolation results

\subsection{Reinitialization}
Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of CFL number, potentially lead degrades the signed distance property of level-set function. It is thus important to reinitialize the level-set every few iterations, especially because the quality of generated grid heavily depends on the signed distance property. To achieve this property we solve the pseudo-time transient equation \eqref{eq:reinitialization} using the discretization scheme detailed in \cite{Min;Gibou:07:A-second-order-accur}. For completeness, we briefly review the scheme here. First equation \eqref{eq:reinitialization} is written in the following semi-discrete form:
\be
\frac{\text{d} \phi}{\text{d} \tau} + S(\phi_0) \left(\mathcal{H}_G(D^+_i \phi, D^-_i \phi) - 1 \right) = 0,
\label{eq:semidiscrete_reinit}
\ee
where $D^+_i \phi$ and $D^-_i \phi$ are the forward and backward derivatives in the $x_i$ direction and $\mathcal{H}_G$ is the Godunov Hamiltonian defined as:\
\ben
\mathcal{H}_G(a_i, b_i) = 
\left\{
\begin{array}{lcr}
	\sqrt{\sum_i \max\left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
	\\
	\sqrt{\sum_i \max\left(|a^-_i|^2, |b^+_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0)  >  0 
\end{array}
\right.
,
\een
where $a^+ = \max(a, 0)$ and $a^- = \min(a, 0)$. Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme using adaptive time-stepping to accelerate convergence to steady state. Since all computation is grid based, parallel implementation of this scheme is mostly trivial. However, one minor point requires further explanation. As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require computation of second derivatives. To enable overlap between computation and communications when computing second derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique. First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$. Here, boundary points are the collection of all local points that are regarded as ghost point, $G_p$, on all other processors, i.e. $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$. Private points are fined as the collection of all local points that are not a boundary point, i.e. $L_p = P_p\cup B_p$. Algorithm \ref{alg:overlap} illustrates how this labeling can help with computation/computation overlap of an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$. Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.

\begin{algorithm}[htbp]
\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$}\label{euclid}
\begin{algorithmic}[1]
\For {$i:B_p$} \Comment{I -- perform computation on boundary points}
	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$
\EndFor
\State $send\_req \gets \texttt{MPI\_Isend(}\phi^{n+1}_B\texttt{)}$ \Comment{II -- begin updating ghost values}
\State $recv\_req \gets \texttt{MPI\_Irecv(}\phi^{n+1}_G\texttt{)}$
\For {$i:P_p$} \Comment{III -- perform computation on private points}
	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$ 
\EndFor
\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- finish updating ghost values}
\State \Return $\phi^{n+1}$
\end{algorithmic}
\label{alg:overlap}
\end{algorithm}

>>>>

\subsection{Reinitialization}
Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of CFL number, potentially lead degrades the signed distance property of level-set function. It is thus important to reinitialize the level-set every few iterations, especially because the quality of generated grid heavily depends on the signed distance property. To achieve this property we solve the pseudo-time transient equation \eqref{eq:reinitialization} using the discretization scheme detailed in \cite{Min;Gibou:07:A-second-order-accur}. For completeness, we briefly review the scheme here. First equation \eqref{eq:reinitialization} is written in the following semi-discrete form:
\be
\frac{\text{d} \phi}{\text{d} \tau} + S(\phi_0) \left(\mathcal{H}_G(D^+_i \phi, D^-_i \phi) - 1 \right) = 0,
\label{eq:semidiscrete_reinit}
\ee
where $D^+_i \phi$ and $D^-_i \phi$ are the forward and backward derivatives in the $x_i$ direction and $\mathcal{H}_G$ is the Godunov Hamiltonian defined as:\
\ben
\mathcal{H}_G(a_i, b_i) = 
\left\{
\begin{array}{lcr}
	\sqrt{\sum_i \max\left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
	\\
	\sqrt{\sum_i \max\left(|a^-_i|^2, |b^+_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0)  >  0 
\end{array}
\right.
,
\een
where $a^+ = \max(a, 0)$ and $a^- = \min(a, 0)$. Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme using adaptive time-stepping to accelerate convergence to steady state. Since all computation is grid based, parallel implementation of this scheme is mostly trivial. However, one minor point requires further explanation. As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require computation of second derivatives. To enable overlap between computation and communications when computing second derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique. First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$. Here, boundary points are the collection of all local points that are regarded as ghost point, $G_r$, on all other processors, i.e. $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$. Private points are fined as the collection of all local points that are not a boundary point, i.e. $P_p = L_p \setminus B_p$. Algorithm \ref{alg:overlap} illustrates how this labeling can help with computation/computation overlap of an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$. Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.

\begin{algorithm}[htbp]
\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$}\label{euclid}
\begin{algorithmic}[1]
\For {$i:B_p$} \Comment{I -- perform computation on boundary points}
	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$
\EndFor
\State $send\_req \gets \texttt{MPI\_Isend(}\phi^{n+1}_B\texttt{)}$ \Comment{II -- begin updating ghost values}
\State $recv\_req \gets \texttt{MPI\_Irecv(}\phi^{n+1}_G\texttt{)}$
\For {$i:P_p$} \Comment{III -- perform computation on private points}
	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$ 
\EndFor
\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
\State \Return $\phi^{n+1}$
\end{algorithmic}
\label{alg:overlap}
\end{algorithm}

### 2014-11-13 paper/tex/scaling_results.tex
added interpolation results

%!TEX root = draft.tex
\section{scalability and accuracy}

\subsection{Interpolation}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_interpolation_all.pdf}
\caption{Scalability of the interpolation procedure. blue: alpha=0.05, purple alpha=0.50, red alpha=0.95 (alpha = percent of non-local points)}
\end{center}
\end{figure}

\subsection{Semi lagrangian advection}

\subsection{Reinitialization} \label{section::scaling_reinitialization}

The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space. The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment. This information can either be computed on the fly or stored in a buffer. Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load. However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation. If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.

The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability. Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
\caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
\end{center}
\end{figure}

\subsection{Poisson solver}

>>>>

%!TEX root = draft.tex
\section{Scaling} \label{sec:scaling}

\subsection{Interpolation}
In this section we show results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.

\begin{figure}[htbp]
	\begin{center}
		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
		\\
		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
	\end{center}
\end{figure}

% \subsection{Semi lagrangian advection}

% \subsection{Reinitialization} \label{section::scaling_reinitialization}

% The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space. The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment. This information can either be computed on the fly or stored in a buffer. Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load. However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation. If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.

% The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability. Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.

% \begin{figure}[ht!]
% \begin{center}
% \includegraphics[width=.7\textwidth]{figures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
% \caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
% \end{center}
% \end{figure}

% \subsection{Poisson solver}
